Fluid-mechanical erosion creates anisotropic porous media

### 2022-05-25 docs/PhysicsPaper/StokesErosion.tex
NM

%-----------------------------------------------------------------------------------------------%
\subsection{Resistivity and permeability}
Darcy's law is given by
\begin{equation}
\bq = - \bvec{K} \grad \phi
\end{equation}
where $\bq$ is the specific discharge, $\phi = z + p/\gamma$ is the hydraulic head, and $\gamma = \rho g$. The rank-2 tensor $\bvec{K}$ is the {\em conductivity}, which is related to the {\em permeability} $\bvec{k}$ via $\bvec{K} = \bvec{k} \gamma/\mu$, where $\mu$ is the viscosity of the fluid. The units of conductivity are length-per-time (i.e.~velocity) and the units of permeability are length squared.
Perhaps the most standard parameter to characterize the porous medium is the permeability. We eliminate gravity from consideration to get
\begin{equation}
\bq = - \bvec{k} \grad p / \mu
\end{equation}
Henceforth, we set $\mu = 1$ by the non-dimensionalization.

>>>>

With the numerical methods in place, we now discuss how to measure the permeability and other porous-media properties of the configurations generated by fluid-mechanically induced erosion.

%-----------------------------------------------------------------------------------------------%
\subsection{Permeability}

While the Stokes equations \eqref{eqn:StokesEq} provides a microscopic description of the detailed flow field $\uu$ penetrating the complex configuration of erodable bodies, a coarse-grained description can be obtained by treating the collection of bodies as a single porous-medium and homogenizing the flow-field through Darcy's law,
\begin{equation}
\bq = - \bvec{k} \grad p / \mu
\end{equation}
Here, $p$ and $\mu$ represent the pressure field and fluid viscosity as before, with $\mu = 1$ by the non-dimensionalization. Meanwhile, $\bq$ represents the {\em specific discharge}, which is the volume of water that flowing through a unit cross sectional area of porous media per unit time; $\bq$ relates to the (interstitial) velocity $\uu$, by integrating $\uu$ over a sufficiently small control region and dividing by the total volume (including both fluid and solid) of the region. The parameter $\bvec{k}$ represents the {\em permeability} of the porous medium, which generally takes the form of a rank-2 tensor to permit different propensities to flow in different directions, i.e.~medium anisotropy. The permeability may vary in space and will certainly vary in time as the bodies erode.

Consider, first for simplicity, an isotropic porous medium with scalar permeability $k$.

% Notes
% Specific discharge: It represents the volume of water that flows through a unit cross sectional area of porous media per unit time. The units of permeability are length squared.
% Reference: https://books.gw-project.org/hydrogeologic-properties-of-earth-materials-and-principles-of-groundwater-flow/chapter/darcys-law/

\vsp{2} \nick{To here in editing} \vsp{2}
